\documentclass[sigconf,review,anonymous]{acmart}
%\documentclass[sigconf,review]{acmart}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{textcomp}
\usepackage{booktabs,subfig,graphicx,hyperref,multirow,longtable}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{threeparttable}

\newcommand{\td}{\textcolor{red}{TODO: }}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
    
% These commands are for a PROCEEDINGS abstract or paper.
\copyrightyear{2019}
\acmYear{2019}
\setcopyright{acmlicensed}
\acmConference[SC'19]{SC'19: The International Conference for High Performance Computing, Networking, Storage and Analysis}{November 17–-22, 2019}{Denver, Colorado}
\acmBooktitle{SC'19: The International Conference for High Performance Computing, Networking, Storage and Analysis, November 17–22, 2019, Denver, Colorado}
\acmPrice{15.00}
\acmDOI{10.1145/1122445.1122456}
\acmISBN{978-1-4503-9999-9/18/06}


% Submission ID. 
%\acmSubmissionID{123-A56-BU3}
\begin{document}

%
% The "title" command has an optional parameter, allowing the author to define a "short title" to be used in page headers.
\title{Feluca: A High-Performance Graph Coloring Algorithm on GPU}

% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands used to denote shared contribution to the research.
\author{Zhigao Zheng$^+$, Xuanhua Shi$^+$, Ligang He$^{\S}$, Hai Jin$^+$, Shuo Wei$^+$, Hulin Dai$^+$, Xuan Peng$^+$}
%\authornote{Both authors contributed equally to this research.}
%\authornotemark[1]
\affiliation{%
  \institution{$^+$National Engineering Research Center for Big Data Technology and System / Services Computing Technology and System Lab}
	\institution{Huazhong University of Science and Technology, China}	
	\institution{$^{\S}$Department of Computer Science}
	\institution{University of Warwick, United Kingdom}
	%\institution{$^{++}$Department of Electrical and Computer Engineering}
	%\institution{George Washington University, USA}	
  %\streetaddress{P.O. Box 1212}
	%\city{*Wuhan}
  %\state{*Hubei}
  %\postcode{430074}
	%\country{China}
}
\email{{zhengzhigao,xhshi}@hust.edu.cn; ligang.he@warwick.ac.uk; {hjin,weishuo,hulindai,piecesix}@hust.edu.cn}

\renewcommand{\shortauthors}{Zhigao and Xuanhua, et al.}

\begin{abstract}
Graph coloring is a task of partitioning the nodes in a graph in such a way that two adjacent nodes reside in two different partitions while the number of total partitions is minimized. All the nodes in the same partition are regarded as having the same color. Graph coloring has been widely used in several applications including air traffic management, parallel computing resource allocation and scheduling, community detection. However, there still exist the great challenges in coloring a large-scale graph on GPU. 
First, the long-tail problem exists in the recursion algorithm because the conflicting (i.e., different threads assign the adjacent nodes to the same color) becomes more likely to occur as the number of iterations increases. Second, it is hard to parallelize the sequential spread algorithm because in the coloring procedure the coloring behaviour in the current iteration is dependent on the coloring results of the previous iteration. Third, the atomic operation is widely used on GPU to maintain the color list, which can greatly reduce the efficiency of GPU threads.

In this paper, we propose a two-stage high-performance graph coloring algorithm, called \textbf{Feluca}, aiming to address the above challenges.
Feluca combines the recursion-based method with the sequential spread-based method. In the first stage, Feluca uses a recursive routine to color a majority of vertices in the graph. Then, it switches to the sequential spread method to color the remaining vertices in order to avoid the conflicts of the recursive algorithm.
Moreover, the following techniques are proposed to further improve the graph coloring performance. \romannumeral1) A new method is proposed to eliminate the cycles in the graph; \romannumeral2) a top-down scheme is developed to avoid the atomic operation originally required for color selection; and \romannumeral3) a novel coloring  paradigm is designed to improve the degree of parallelism for the sequential spread part. All these newly developed techniques, together with further GPU-specific optimizations such as coalesced memory access, comprise an efficient parallel graph coloring solution in Feluca. We have conducted extensive experiments on NVIDIA K20m GPU. The results show that Feluca can achieve 3.61--125.81$\times$ speedup over the greedy coloring algorithm and 18.58--71.07$\times$ speedup over the existing Jones-Plassman-Luby (JPL) coloring algorithm. 

\end{abstract}

% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
% Please copy and paste the code instead of the example below.
\begin{CCSXML}
<ccs2012>
 <concept>
	<concept_id>10010147.10010169.10010170.10010174</concept_id>
	<concept_desc>Computing methodologies~Massively parallel algorithms</concept_desc>
	<concept_significance>500</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Massively parallel algorithms}

\keywords{Graph Coloring, GPGPU, Parallelism, Color-centric Paradigm, Pipeline}

\maketitle

\input{sec_1_Introduction}
\input{sec_2_motivation}
\input{sec_3_design}
\input{sec_4_feluca}
\input{sec_5_experiments}
\input{sec_6_related_work}
\input{sec_7_conclusion}

\begin{acks}
This work is partly supported by National Key R\&D Program of China (No. 2017YFC0803700), two grants 
from the National Science Foundation of China (No. 61772218 and No. 61433019) and a grant 
from the Outstanding Youth Foundation of Hubei Province (No. 2016CFA032).
\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{acmart}

% If your work has an appendix, this is the place to put it.
%\appendix
\end{document}
